Appendix: Trends in age-adjusted mortality rates for blacks and whites by cause and sex, 1969-2013, United States

### 2017-04-06 Manuscript/Appendix-agestd-mort-rates.Rmd
use state abbrevs

mortality.rates <- reorder.as.map(mortality.rates, "state")
mortality.rates.diff <- reorder.as.map(mortality.rates.diff, "state")
```

@@

  a <- ggplot(subset.data, aes(x = year, y = rate.per.100k_mean)) + 
              geom_ribbon(aes(ymin = rate.per.100k_lcl, ymax = rate.per.100k_ucl, fill = Race), col = NA, alpha = 0.5) +
              geom_line(aes(col = Race)) +
              facet_wrap( ~ state.map.order, ncol = 11, drop = F) +
              xlab(" ") + 
              ylab("Age-standardized mortality rate (per 100,000)") + 
              geom_hline(yintercept = 0, col = "black", lwd = 0.5) +
              #scale_y_continuous(limits = lim.y) + 
              ggtitle(paste0("Age-adjusted mortality (per 100,000) in ", sex1, " for ", COD1)) +
              theme_classic(base_size = 10) +
              theme(axis.text.x = element_blank(),
                    strip.background=element_blank(),
                    axis.line=element_blank(),
                    axis.ticks=element_blank())

  b <- ggplot(subset.data2, aes(x = year, y = rate.difference_mean)) + 
              geom_line() +
              geom_ribbon(aes(ymin = rate.difference_LCL, ymax = rate.difference_UCL), col = NA, alpha = 0.5) +
              facet_wrap( ~ state.map.order, ncol = 11, drop = F) +
              xlab(" ") + 
              ylab("Excess mortality among blacks (per 100,000)") +
              #scale_y_continuous(breaks = 0) + #limits = lim.y2
              geom_hline(yintercept = 0, col = "black", lwd = 0.5) +
              #geom_vline(xintercept = 1969, col = "black", lwd = 0.5) +
              ggtitle(paste0("Excess mortality among blacks (per 100,000) in ", sex1, " for ", COD1)) +
              theme_classic(base_size = 10) +
              theme(panel.grid.minor = element_blank(),
                    axis.text.x = element_blank(),
                    strip.background=element_blank(),
                    axis.line=element_blank(),
                    axis.ticks=element_blank())

@@

```{r print-em, fig.width = 9, fig.height = 9, echo = F}
ggplotly(Cardiovascular.Male.mort.plot[[1]]) %>% layout(xaxis = list(title = "Year (1969-2013)"), showlegend = F) 
ggplotly(Cardiovascular.Male.mort.plot[[2]]) %>% layout(xaxis = list(title = "Year (1969-2013)"))

@@

```{r print-em-2, fig.width = 9, fig.height = 9, echo = F}
ggplotly(Cardiovascular.Female.mort.plot[[1]]) %>% layout(xaxis = list(title = "Year (1969-2013)"), showlegend = F) 
ggplotly(Cardiovascular.Female.mort.plot[[2]]) %>% layout(xaxis = list(title = "Year (1969-2013)"))
```

>>>>

#DC state abbreviation wasn't accurately defined. Adding it here.
levels(mortality.rates$stabbrs) <- c(levels(mortality.rates$stabbrs), "DC")
levels(mortality.rates.diff$stabbrs) <- c(levels(mortality.rates.diff$stabbrs), "DC")
mortality.rates$stabbrs[mortality.rates$state == "Washington DC"] <- "DC"
mortality.rates.diff$stabbrs[mortality.rates.diff$state == "Washington DC"] <- "DC"

mortality.rates <- reorder.as.map(mortality.rates, "state", "stabbrs")
mortality.rates.diff <- reorder.as.map(mortality.rates.diff, "state", "stabbrs")
```

@@

  a <- ggplot(subset.data, aes(x = year, y = rate.per.100k_mean)) + 
              geom_ribbon(aes(ymin = rate.per.100k_lcl, ymax = rate.per.100k_ucl, fill = Race), col = NA, alpha = 0.5) +
              geom_line(aes(col = Race)) +
              facet_wrap( ~ stabbrs.map.order, ncol = 11, drop = F) +
              xlab(" ") + 
              ylab("Age-standardized mortality rate (per 100,000)") + 
              geom_hline(aes(yintercept = 0)) + geom_vline(aes(xintercept = 1969)) +  
              ggtitle(paste0("Age-adjusted mortality (per 100,000) in ", sex1, " for ", COD1)) +
              theme_classic(base_size = 10) +
              theme(axis.text.x = element_blank(),
                    strip.background=element_blank(),
                    axis.line=element_blank(),
                    axis.ticks=element_blank())

  b <- ggplot(subset.data2, aes(x = year, y = rate.difference_mean)) +
           # geom_hline(aes(yintercept = -300), lwd = 0.5, col = "grey") +
           # geom_hline(aes(yintercept = -200), lwd = 0.5, col = "grey") +
           # geom_hline(aes(yintercept = -100), lwd = 0.5, col = "grey") +
           # geom_hline(aes(yintercept = 0), lwd = 0.5, col = "grey") +
           # geom_hline(aes(yintercept = 200), lwd = 0.5, col = "grey") +
           # geom_hline(aes(yintercept = 100), lwd = 0.5, col = "grey") +
              geom_line() +
              geom_ribbon(aes(ymin = rate.difference_LCL, ymax = rate.difference_UCL), col = NA, alpha = 0.5) +
              facet_wrap( ~ stabbrs.map.order, ncol = 11, drop = F) +
              xlab(" ") + 
              ylab("Excess mortality among blacks (per 100,000)") +
              geom_hline(aes(yintercept = 0)) + geom_vline(aes(xintercept = 1969)) +  
              ggtitle(paste0("Excess mortality among blacks (per 100,000) in ", sex1, " for ", COD1)) +
              theme_classic(base_size = 10) +
              theme(panel.grid.minor = element_blank(),
                    axis.text.x = element_blank(),
                    strip.background=element_blank(),
                    axis.line=element_blank(),
                    axis.ticks=element_blank())

@@

```{r print-em, fig.width = 9, fig.height = 9, echo = F}
ggplotly(Cardiovascular.Male.mort.plot[[1]]) %>% 
  layout(xaxis = list(title = "Year (1969-2013)"), showlegend = F)

ggplotly(Cardiovascular.Male.mort.plot[[2]]) %>% 
  layout(xaxis = list(title = "Year (1969-2013)"))

@@

```{r print-em-2, fig.width = 9, fig.height = 9, echo = F}
ggplotly(Cardiovascular.Female.mort.plot[[1]]) %>% 
  layout(xaxis = list(title = "Year (1969-2013)"), showlegend = F)

ggplotly(Cardiovascular.Female.mort.plot[[2]]) %>% 
  layout(xaxis = list(title = "Year (1969-2013)"))
```
